---
title: "p8105_mtp_zz2780"
author: "Lydia Zhang"
date: "2022-10-20"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(tibble)
library(janitor)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .5,
  out.width = "90%"
)
library(patchwork)
```


Word Count: `r wordcountaddin::word_count("p8105_mtp_zz2780.Rmd")`

Problem 1 - Data

The raw data has `r ncol(data)` variables and `r nrow(data)` observations. Under the columns Raison, Simone, Gagne are the weights recorded for each dog on different dates. And std represents the standard object's weight. Some dates contain notes recorded that day. The goal of this report this to tidy the data set, conduct descriptive analysis, and produce a graphic output for the distribution of each dog's weight and their trends over time.

For the first data frame that contains weights of each dog and the dates recorded, I first imported the table from the excel sheet and omitted the Notes section. After I cleaned the columns' names, I realized that variables recording the weights contained two values, one in lb and one in oz, separated by space. After separating the two numbers in one variable into two and converting them into numeric characters, I combine the two columns with same unit in lb for Raison, Simone, Gagne, and the standard object, which were rounded to two decimal places. Then I dropped all the intermediate variables and only kept weights in lb for Raison, Simone, Gagne, their std, and the dates recorded. 

Since the column names of dogs and standard object is also a variable, I pivoted the table so that there is a variable called Name, including values Raisin, Simone, Gagne, and std, and the second variable Weight with each object's weight recorded on that date. After dropping all NA values, finally I converted the numeric values of dates to modern date format. The first data frame is exported with the name "dataframe1.csv" in the data folder along with the raw data.

```{r}
data=read_excel("data/dogweights_07may2022.xlsx", range="B2:F363" )
data1<-data%>%
  janitor::clean_names() %>%
  separate(col=raisin, into=c("raison_lb", "raison_oz"), sep=" ")%>%
  separate(col=simone, into=c("simone_lb", "simone_oz"), sep=" ")%>%
  separate(col=gagne, into=c("gagne_lb", "gagne_oz"), sep=" ")%>%
  separate(col=std, into=c("std_lb", "std_oz"), sep=" ")%>%
  transform(
            raison_lb=as.numeric(raison_lb),
            raison_oz=as.numeric(raison_oz),
            simone_lb=as.numeric(simone_lb),
            simone_oz=as.numeric(simone_oz),
            gagne_lb=as.numeric(gagne_lb),
            gagne_oz=as.numeric(gagne_oz),
            std_lb=as.numeric(std_lb),
            std_oz=as.numeric(std_oz)
  )%>%
  mutate(
    raison=round(raison_lb+0.0625*raison_oz,2),
    simone=round(simone_lb+0.0625*simone_oz,2),
    gagne=round(gagne_lb+0.0625*gagne_oz,2),
    std=round(std_lb+0.0625*std_oz,2)
  )%>%
  select(date, raison, simone, gagne, std)%>%
  pivot_longer(
    raison:std,
    names_to="name",
    values_to="weight"
  )%>%
  na.omit()

data1$date=excel_numeric_to_date(as.numeric(data1$date), date_system="modern")

write.csv(data1, "data/dataframe1.csv", row.names=FALSE)
```

The second data frame was selected from the same excel sheet. After cleaning the name, I copied the column of notes and gave the new column name "Notes". I also changed the numeric date format to modern format. And I only selected the variables of notes written as text and dates recorded and removed all NA values. I exported the second data frame as "dataframe2.csv" to the data folder along with the raw data and dataframe1. 
```{r}
data2=read_xlsx("data/dogweights_07may2022.xlsx", range="B2:G363" )%>%
  janitor::clean_names()%>%
  mutate(Notes=x6,
         date=excel_numeric_to_date(as.numeric(date), date_system="modern")
         )%>%
  select(date, Notes)%>%
  na.omit()

write.csv(data2, "data/dataframe2.csv", row.names=FALSE)
```

Problem 2 - EDA
The resulting tidy data set contains `r nrow(data1)` observations and `r ncol(data1)` observations. Variables are recorded date, dog's name, and corresponding weight.

```{r}
data1%>%
  count(date, name="n_obs")%>%
  na.omit()
```

356 distinct dates are include in the dataset. 

The table below recorded the number of observations for each dog, and their average weight and standard deviation. Gagne has 119 recorded weights, with mean of 10.29lb, and sd of 0.71lb. Raison has 352 recorded weights, with an average of 17.78lb, and sd of 0.94lb. Simone has 333 recorded weights, with an average of 9.79 and 0.92 standard deviation. The standard object has 45 recorded weights, with an average of 20.98lb and standard deviation of 0.01.

```{r}
table1<-data1%>%
  group_by(name)%>%
  na.omit()%>%
  summarize(
    n_obs=n(),
    avg_weight=round(mean(weight),2),
    sd_weight=round(sd(weight),2)
  )
table1
```


Problem 3 - Visualization

```{r}
plot1<-data1%>%
  ggplot(aes(x=name, y=weight))+
  geom_boxplot()+
  labs(title="Distrbution of Weight")

plot2<-data1%>%
  na.omit()%>%
   ggplot(aes(x=date,y=weight))+
  geom_line(aes(color=name))+
  labs(title="Trend of Weight Over Time")

plot<-plot1 | plot2
plot

ggsave("results/Weight_plot.pdf",plot, width=13, height=7)
```






